MCC Model Predictions

### 2018-06-28 stats/ex3_comparisons.Rmd
updating exp variable priors

library(tidyverse)
library(knitr)
library(ggthemes)
library(langcog)
library(rwebppl)
library(matrixStats)
library(coda)

@@

pred2 <- predictions %>%
  group_by(Model,Experiment,Change,Alignment)%>%
  multi_boot_standard(col = "corr")

@@

@@

@@

@@

@@

```{r}
## joining noisy model predictions with other predictions
no.adj.models <- pd %>%
  filter(Model != "Full model", Model != "Prior only")%>%
  mutate(Type = ifelse(Model == "Full model","RSA",ifelse(Model == "Data","Data",ifelse(Model == "No prior","No Prior","Prior only"))),
         Adjustment = "None")

@@

p.com$Data = rep(pd$mean[pd$Model=="Data"],10)
p.com$Data_ci_low = rep(pd$ci_lower[pd$Model=="Data"],10)
p.com$Data_ci_up = rep(pd$ci_upper[pd$Model=="Data"],10)

@@

ggplot(data = plot.model,aes(x = mean, y = Data, col = Model)) +
  geom_abline(intercept = 0, slope = 1, lty = 2, alpha = 0.3)+
  geom_point()+
  geom_errorbar(aes(ymin = Data_ci_low, ymax = Data_ci_up),width = 0)+
  geom_errorbarh(aes(xmin = ci_lower, xmax = ci_upper), height = 0)+
  facet_grid(Type~Adjustment)+
  coord_fixed()+
  xlim(0,1)+ylim(0,1)+
  xlab("Model")+
  stat_cor(method = "pearson", label.x = 0.4, label.y = 0.9)+
  geom_smooth(method = "lm", se = F, col = "black", size =0.4)+
  theme_few() + 
  guides(col = F)+
  scale_colour_solarized()

>>>>

library(tidyverse)
library(knitr)
library(ggthemes)
library(langcog)
library(rwebppl)
library(matrixStats)
library(coda)
library(ggpubr)

@@

pred <- predictions %>%
  group_by(Model,Experiment,Change,Alignment)%>%
  multi_boot_standard(col = "corr")

@@

da2 <- data %>%
  filter(trial_type=="test")%>%
  mutate(Change = ifelse(change =="same", "Same speaker", "Different speaker"),
         Alignment = ifelse(alignment == "congruent", "Congruent", "Incongruent"),
         Experiment = ifelse(experiment == "pref_inf", "Preference", "Novelty"),
         Model = "Data") %>%
  group_by(Model,Experiment,Change,Alignment)%>%
  summarize(k = sum(correct_inf), n = n())%>%
  ungroup() %>%
  mutate(a = 1 + k,
         b = 1 + n - k,
         ci_lower  = qbeta(.025, a, b),
         ci_upper = qbeta(.975, a, b),
         mean = (a-1)/(a+b-2))%>%
  select(-a,-b,-n,-k)

@@

pd2 <- bind_rows(pred,da2)

@@

noisy.rsa.params <- bda %>%
  filter(Parameter %in% c("noise", "priorExponent"))  %>%
  group_by(Parameter) %>%
  summarise(mean = estimate_mode(value), ci_lower = hdi_lower(value), ci_upper = hdi_upper(value))

noisy.rsa.params100k <- bda2 %>%
  filter(Parameter %in% c("noise", "priorExponent"))  %>%
  group_by(Parameter) %>%
  summarise(mean = estimate_mode(value), ci_lower = hdi_lower(value), ci_upper = hdi_upper(value))

@@

noisy.prior.params <- bda1 %>%
  filter(Parameter %in% c("noise", "priorExponent"))  %>%
  group_by(Parameter) %>%
  summarise(mean = estimate_mode(value), ci_lower = hdi_lower(value), ci_upper = hdi_upper(value))

@@

```{r}
## joining noisy model predictions with other predictions
no.adj.models <- pd2 %>%
  filter(Model != "Full model", Model != "Prior only")%>%
  mutate(Type = ifelse(Model == "Full model","RSA",ifelse(Model == "Data","Data",ifelse(Model == "No prior","No Prior","Prior only"))),
         Adjustment = "None")

@@

p.com$Data = rep(pd2$mean[pd2$Model=="Data"],10)
p.com$Data_ci_low = rep(pd2$ci_lower[pd2$Model=="Data"],10)
p.com$Data_ci_up = rep(pd2$ci_upper[pd2$Model=="Data"],10)

@@

ggplot(data = plot.model,aes(x = mean, y = Data, col = Model)) +
  geom_abline(intercept = 0, slope = 1, lty = 2, alpha = 0.3)+
  geom_point()+
  geom_errorbar(aes(ymin = Data_ci_low, ymax = Data_ci_up),width = 0)+
  geom_errorbarh(aes(xmin = ci_lower, xmax = ci_upper), height = 0)+
  facet_grid(Type~Adjustment)+
  coord_fixed()+
  xlim(0,1)+ylim(0,1)+
  xlab("Model")+
  stat_cor(method = "pearson", label.x = 0.3, label.y = 0.9)+
  geom_smooth(method = "lm", se = F, col = "black", size =0.4)+
  theme_few() + 
  guides(col = F)+
  scale_colour_solarized()

### 2018-06-28 stats/mcc_pref.kids.Rmd
updating exp variable priors

# children who need to be dropped
drop <- c("Dryrun","180426_3_pref","180502_1_pref","180503_1_pref","180503_2_pref","180504_1_pref","180511_2_pref","180511_1_pref")

@@

@@

# maximal converging model
lm <- glmer(correct ~ age + change + (age * change |id) + (age * change |choiceAgent), 
              data = data_cg, family = binomial, 
              control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))

>>>>

# children who need to be dropped
drop <- c("Dryrun",
"180426_3_pref",
"180427_2_pref",
"180427_3_pref",
"180427_4_pref",
"180502_1_pref",
"180503_1_pref",
"180503_2_pref",
"180504_1_pref",
"180511_1_pref",
"180511_2_pref")

@@

unique(d$id)
```

@@

# maximal converging model
lm <- glmer(correct ~ age * change + (age * change |id) + (age * change |choiceAgent), 
              data = data_cg, family = binomial, 
              control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))

### 2018-06-28 stats/model_predictions_variable_priors.Rmd
variable prior exp and model predictions
updating exp variable priors

```{r predictions rsa noise, include = FALSE}
predictions.rsa.noise <-'
var allConditions = dataFromR.newConditions
var allData = dataFromR.predData
var allPriorProbs = dataFromR.predPriorProbs

@@

mediumPriorProbs$priorProbs = list(
    c(0.1465517, 0.1465517, 0.8534483), 
    c(0.8534483, 0.8534483, 0.1465517),
    c(0.4137931, 0.4137931, 0.5862069),
    c(0.5862069, 0.5862069, 0.4137931),
    c(0.2983871, 0.2983871, 0.7016129),
    c(0.7016129, 0.7016129, 0.2983871),
    c(0.4137931, 0.4137931, 0.5862069),
    c(0.5862069, 0.5862069, 0.4137931))

@@

@@

@@

```{r}

@@

>>>>

mediumPriorProbs$priorProbs = list(
    c(0.1465517, 0.1465517, 0.8534483), 
    c(0.8534483, 0.8534483, 0.1465517),
    c(0.4137931, 0.4137931, 0.5862069),
    c(0.5862069, 0.5862069, 0.4137931),
    c(0.2983871, 0.2983871, 0.7016129),
    c(0.7016129, 0.7016129, 0.2983871),
    c(0.4274194, 0.4274194, 0.5725806),
    c(0.5725806, 0.5725806, 0.4274194))

```{r model rsa noise, include = FALSE}
predictions.rsa.noise <-'
var allConditions = dataFromR.newConditions
var allData = dataFromR.predData
var allPriorProbs = dataFromR.predPriorProbs

@@

@@

#bda.pred.noise2 has 100k samples

@@

noisy.rsa.params.pred <- bda.pred.noise %>%
  filter(Parameter %in% c("noise", "priorExponent"))  %>%
  group_by(Parameter) %>%
  summarise(mean = estimate_mode(value), ci_lower = hdi_lower(value), ci_upper = hdi_upper(value))

noisy.rsa.params.pred100k <- bda.pred.noise2 %>%
  filter(Parameter %in% c("noise", "priorExponent"))  %>%
  group_by(Parameter) %>%
  summarise(mean = estimate_mode(value), ci_lower = hdi_lower(value), ci_upper = hdi_upper(value))

@@

```{r plot rsa noise pred}

@@

```{r model prior only noise , include = FALSE}
predictions.prior.noise <-'
var allConditions = dataFromR.newConditions
var allData = dataFromR.predData
var allPriorProbs = dataFromR.predPriorProbs

  var modelPredictions = Infer({method: "enumerate", model: function(){
      var obj = sample( Categorical({vs: all_objects, ps: conditionSpecificPriors[0].priorProbs}));
      return obj.shape == "circle" ? 1 : 0
  }})

   var noisyModelPredictions = addNoise(modelPredictions, noise)

   map(function(d){ 
     // display(JSON.stringify(noisyModelPredictions) + JSON.stringify(d.correct_inf))
      observe(noisyModelPredictions, d.correct_inf)
    }, conditionSpecificData)

  return  [conditionInfo.Experiment + "_" + conditionInfo.Change + "_" +  conditionInfo.Alignment, Math.exp(noisyModelPredictions.score(1))]

    var newModelPredictions = Infer({method: "enumerate", model: function(){
      var obj = sample( Categorical({vs: all_objects, ps: newSpecificPriors[0].priorProbs}));
      return obj.shape == "circle" ? 1 : 0
  }})

  var newNoisyModelPredictions = addNoise(newModelPredictions, noise)

```{r predictions prior only noise, include = FALSE}
bda.pred.prior.noise <- webppl(
  program_code = paste(rsaUtils, rsaModel2, predictions.prior.noise, sep='\n'),
  data = list(allData = data, newConditions = newConditions, predPriorProbs = predPriorProbs, predData = predData), 
  data_var = "dataFromR",
  model_var = "model",
  inference_opts = list(method = "MCMC", samples = 10000, burn = 2000, verbose = T)
)

## distribution of noise
posterior.prior.noise.pred <- bda.pred.prior.noise %>%
  filter(Parameter %in% c("noise", "priorExponent"))

ggplot(posterior.prior.noise.pred, aes(x = value))+
  geom_histogram()+
  facet_wrap(~Parameter, scales = 'free')

noisy.prior.params.pred <- bda.pred.prior.noise %>%
  filter(Parameter %in% c("noise", "priorExponent"))  %>%
  group_by(Parameter) %>%
  summarise(mean = estimate_mode(value), ci_lower = hdi_lower(value), ci_upper = hdi_upper(value))

prior.noise.pred <- bda.pred.prior.noise %>%
  filter(!(Parameter %in% c("noise", "priorExponent")))  %>%
  separate(Parameter, into = c("Experiment", "Change", "Alignment", "Prior"), sep="_")%>%
  mutate(Model="Noisy Prior only Model",
         Adjustment = "Noise",
         Type = "Prior only",
         Prior = relevel(as.factor(Prior), ref = "Weak"))%>%
  group_by(Adjustment,Model,Prior,Type,Experiment,Change,Alignment) %>%
  summarise(mean = estimate_mode(value), ci_lower = hdi_lower(value), ci_upper = hdi_upper(value))

```{r plot prior only noise regularized}

ggplot(prior.noise.pred, 
       aes(x = Change, y = mean, fill = Alignment)) +
  geom_bar(stat="identity", position = position_dodge(), color = 'black') + 
  geom_hline(yintercept = 0.5, lty=2)+
  geom_linerange(aes(ymin = ci_lower, ymax = ci_upper),position = position_dodge(width = 0.9))+
  labs(x=" ",y="Proportion Choosing More Informative")+
  facet_grid(Experiment ~ Prior)+
  theme_few(base_size = 12) + 
  theme(axis.text.x=element_text(angle = 45, vjust = 1, hjust = 1))+
  ylim(0,1)
```

```{r}
pred.plot <- bind_rows(prior.noise.pred,noisy.rsa.pred)

ggplot(pred.plot, 
       aes(x = Change, y = mean, fill = Alignment)) +
  geom_bar(stat="identity", position = position_dodge(), color = 'black') + 
  geom_hline(yintercept = 0.5, lty=2)+
  geom_linerange(aes(ymin = ci_lower, ymax = ci_upper),position = position_dodge(width = 0.9))+
  labs(x=" ",y="Proportion Choosing More Informative")+
  facet_grid(Experiment ~ Prior*Type)+
  theme_few(base_size = 12) + 
  theme(axis.text.x=element_text(angle = 45, vjust = 1, hjust = 1))+
  ylim(0,1)
